# Reproducible Research: Peer Assessment 1
**Author:** Florian Knaus
**Last changed:** May 18, 2014
## Introduction
This document is for the Peer Assessment 1 of the Coursera Reproducible Research course by Roger Peng.


## Loading and preprocessing the data
First we set the working directory:
```{r}
setwd("C:/Users/Florian/Documents/Development/git/RepData_PeerAssessment1");
```
Then we read the file "activity.csv" and format the date column
```{r}
activity <- read.table(file="activity.csv", 
                       header=TRUE,
                       sep=",",
                       colClasses=c("numeric","character","numeric"));
activity$date <- as.Date(activity$date, "%Y-%m-%d");
```


## What is mean total number of steps taken per day?
We create a histogram and use functions mean and median on column 'steps' of data-frame 'activity' and remove any possible 'NA' values.
```{r}
hist(activity$steps,
     main="Histogram of Steps", 
     xlab="Steps");
mean(activity$steps, 
     na.rm=TRUE);
median(activity$steps, 
       na.rm=TRUE);
```


## What is the average daily activity pattern?

```{r}
with(activity, plot(x=date, y=steps, type="l"));
with(activity, plot(x=date, y=interval, type="l"));
```



## Imputing missing values
```{r}

```



## Are there differences in activity patterns between weekdays and weekends?
```{r}

```

